A Tale of Two Transects

### 2020-01-20 writing/transect_paper.Rmd

## Clustering

For comparison with clustering-based OTU methods, ASVs were clustered at 97% similarity using VSEARCH v2.9.1 [@rognes2016].

## Alignment and phylogenetic inference

@@

Taxonomic assignments of ASVs which were represented in the long amplicon dataset were refined using the phylogenetic tree.
A taxon at a particular rank was assigned to a node and all its descendents if that taxon was consistent with the reference-based taxonomic assignments for each of the descendents.
A taxon assignment was considered to be consistent if at least one algorithm assigned that taxon at greater than 50% confidence, or if no algorithm successfully classified the sequence at greater than 50% confidence.
The result of this process was to give a taxonomic assignment to ASVs which were previously unassigned if they were nested within a clade where which was consistently given an assignment, as well as clarify the assignment of ASVs where different algorithms had resulted in different assignments, but only one of these was consistent with the assignments of other ASVs in the same clade.

@@

@@

@@

>>>>

### Clustering

For comparison with clustering-based methods, ASVs were clustered into operational taxonomic units (OTUs) at 97% similarity using VSEARCH v2.9.1 [@rognes2016].

### Alignment and phylogenetic inference

@@

Taxonomic assignments of ASVs which were represented in the long amplicon dataset were refined using the phylogenetic tree (Supplementary Figure \@ref(fig:phylotax)).
A taxon at a particular rank was assigned to a node and all its descendents if that taxon was consistent with the reference-based taxonomic assignments for each of the descendents.
A taxon assignment was considered to be consistent if at least one algorithm assigned that taxon at greater than 50% confidence, or if no algorithm successfully classified the sequence at greater than 50% confidence.
The result of this process was to give a taxonomic assignment to ASVs which were previously unassigned if they were nested within a clade where which was consistently given an assignment, as well as clarify the assignment of ASVs where different algorithms had resulted in different assignments, but only one of these was consistent with the assignments of other ASVs in the same clade.

@@

Long (short) amplicon sequencing with PacBio yielded *X* (*Y*) circular consensus reads, of which *A* (*B*) were successfully demultiplexed.
The 5.8S region was detected in *C* (*D*) reads.
Ion Torrent sequencing yielded *E* demultiplexed reads, with the 5.8S region detected in *F*.

@@

Figure \@ref/fig:pb-vs-it) shows the correspondence between the read count for different ASVs/OTUs

@@

(ref:phylotax) Hypothetical example of phylogenetic refinement of taxonomic assignments.
In the example, a clade includes  seven OTUs (A-F), which have been identified by two methods as belonging to taxa "Tax1" or "Tax2", or are unidentified ("unk"), as shown in the tip labels.
No taxon is assigned at node 1, because one of the child branches (A) is completely unassigned.
No taxon is assigned at node 2, because the assignments at C and F are inconsistent.
Node 3 is assigned to Tax1 because this is consistent with at least one of the assignments for both B and C.
Node 4 (and thus also node 5) is assigned to Tax2 because this is consistent with at least one of the assignments for D and F, and because E is completely unassigned.

```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax)"}
extree <-
  tidytree::as.treedata(ape::read.tree(text = "(A:1,((B:1,C:1):1,((E:1,F:1):1,D:1):1):1);")) %>%
  tidytree::full_join(
    tibble(label = LETTERS[1:6],
           tax1 = c("unk", "Tax1", "Tax2", "Tax2", "unk", "unk"),
           tax2 = c("unk", "Tax2", "Tax2", "Tax1", "unk", "Tax1")),
    by = "label"
  )
ggtree::ggtree(extree) +
  ggtree::geom_tiplab(x = 4, align = 2) +
  xlim(0, 6) +
  ggtree::geom_tiplab(
    aes(color = tax1, label = tax1),
    x = 4.7,
    hjust = 0.5
  ) +
  ggtree::geom_tiplab(
    aes(color = tax2, label = tax2),
    x = 5.7,
    hjust = 0.5
  ) +
  ggtree::geom_tiplab(x = 5.2, label = "/", hjust = 0.5) +
  ggtree::geom_nodelab(aes(label = node - 6), geom = "label") +
  ggtree::geom_hilight(node = 10, fill = "steelblue", extendto = 4) +
  ggtree::geom_hilight(node = 9, fill = "tomato", extendto = 4) +
  scale_color_manual(values = c(Tax1 = "steelblue", Tax2 = "tomato", unk = "grey50")) +
  scale_y_reverse()
```
